---
title: "ARIMA_BIke_Share"
author: "Aparajit"
date: "5/28/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library("forecast","tseries","ggplot2")
```
# Time Series forecasting on UCI Bike Share dataset using Box-Jenkins Method

```{r}
bike_data= read.csv('/Users/aparajit/Desktop/PROJECTS/R/ARIMA/Bike-Sharing-Dataset/day.csv',
                    header = TRUE,stringsAsFactors = FALSE)
```
# Data Exploration and outlier treatment using moving averages
```{r}
bike_data$Date=as.Date(bike_data$dteday)
ggplot(bike_data,aes(Date, cnt))+ geom_line()+scale_x_date('month')+
  ylab('Number of bikes')

count_ts=ts(bike_data[,c('cnt')])
bike_data$clean_cnt=tsclean(count_ts)
ggplot(bike_data,aes(Date, clean_cnt))+ geom_line()+scale_x_date('month')+
  ylab('Number of bikes')

bike_data$ma7_cnt= ma(bike_data$clean_cnt,order = 7)
bike_data$ma30_cnt= ma(bike_data$clean_cnt,order = 30)

ggplot()+
  geom_line(data = bike_data,aes(x= Date,y =clean_cnt,colour='Counts' ))+
  geom_line(data = bike_data,aes(x= Date,y =ma7_cnt,colour='Weekly Moving Avg' ))+
  geom_line(data = bike_data,aes(x= Date,y =ma30_cnt,colour='Monthly Moving Avg' ))
```
# Data Decomposition 
```{r}
count_ma=ts(na.omit(bike_data$ma7_cnt), frequency = 30)
decomp=stl(count_ma,s.window = 'periodic')
deseasonal_cnt <- seasadj(decomp)
plot(decomp)
```

# running dickey- fuller test to check stationarity
```{r}
adf.test(count_ma, alternative='stationary')
```
# as a rseult of dicky-fuller test we conclude that series isnt stationary
# as a next step deciding the order of differencing to be applied

```{r}
acf(count_ma)
pacf(count_ma)
```
# attempting differencing of order 1 
```{r}
count_d1 <- diff(deseasonal_cnt,differences = 1)
plot(count_d1)
adf.test(count_d1, alternative='stationary')
```
# series is stationary with d =1
```{r}
acf(count_d1)
pacf(count_d1)
```

# fitting ARIMA model
```{r}
fit <- auto.arima(deseasonal_cnt,seasonal = FALSE)
```
# Model Evaluation

```{r}
tsdisplay(residuals(fit),lag.max = 45,main = '(1,1,1) Residuals')
fit2 <- arima(deseasonal_cnt,order = c(1,1,7))
fit2
tsdisplay(residuals(fit2),lag.max = 15,main = '(1,1,7) Residuals')
```

# Forecasting using this model
```{r}
pred <- forecast(fit2,h = 50)
plot(pred)
```

# including seasonality in the model
```{r}
fit3 <- auto.arima(deseasonal_cnt,seasonal = TRUE)
fit3
pred2 <- forecast(fit3,h=30)
plot(pred2)
tsdisplay(residuals(fit3),lag.max = 15,main = '(2,1,2) Residuals')
```
